### 2018-06-06 preproc/Analysis_exp7_prepro.rmd
add preprocessing code

---
title: "Preprocessing_&_Plot_for_Moral_Categ_Replication"
author: "hcp"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
---
<style type="text/css">

body{ /* Normal  */
   font-family: Times     
   font-size: 12px;
}
td {  /* Table  */
   font-size: 8px;
}
h1 { /* Header 1 */
 font-size: 28px;
}
h2 { /* Header 2 */
 font-size: 22px;
}
h3 { /* Header 3 */
 font-size: 18px;
 color: DarkBlue;
}
code.r{ /* Code block */
  font-size: 10px;
}
pre { /* Code block */
  font-size: 10px
}
</style>

This script is used for quality control and preprocessing of the replication data.

```{r Initializing, include=FALSE,,echo=FALSE,results='hide',warning=FALSE, message=FALSE}
source('Initial.r')

```

```{r loadingData_e7,,echo=FALSE,results='hide',warning=FALSE, message=FALSE}
# set the directories
curDir <- getwd()   # directory for preprocessing
rootDir <- gsub('.{7}$', '', curDir)
rawDir <- paste(curDir,'/data/',sep = '')
traDir <- paste(rootDir,'tradAnal',sep = '')
ddmDir <- paste(rootDir,'hddmMod',sep = '')

# get the file names for files contain "data_exp7_rep_match_"
fNames.m <- list.files(path = rawDir, pattern = "data_exp7_rep_match_*")
# add the directory information to the filenames
fNames.m <- paste(rawDir,fNames.m, sep = '')

# read these files and combine them into one file, get the raw data for matching task
df.L    <- do.call("rbind",lapply(as.list(fNames.m),FUN=function(files){read.table(files, header=TRUE, sep="",stringsAsFactors=F)}))

# do the same to data from categorization, get the raw data for categorization task
fNames.c <- list.files(path = rawDir, pattern = "data_exp7_rep_categ_*")
fNames.c <- paste(rawDir,fNames.c, sep = '')
df.T    <- do.call("rbind",lapply(as.list(fNames.c),FUN=function(files){read.table(files, header=TRUE, sep="",stringsAsFactors=F)}))

# render the data in numeric format for future analysis
cols.num <- c('Sub',"Age","Block","Bin","Trial","RT","ACC")
df.L[cols.num] <- sapply(df.L[cols.num], as.numeric)  
df.T[cols.num] <- sapply(df.T[cols.num], as.numeric)  
df.L <- df.L[!is.na(df.L$ACC),]
df.T <- df.T[!is.na(df.T$ACC),]

# get the independent varialbes
# moral valence
df.L$Morality[grepl("moral", df.L$Shape, fixed=TRUE)]   <- "Good"
df.L$Morality[grepl("immoral", df.L$Shape, fixed=TRUE)] <- "Bad"

# self-referential
df.L$Identity[grepl("Self", df.L$Shape, fixed=TRUE)]    <- "Self"
df.L$Identity[grepl("Other", df.L$Shape, fixed=TRUE)]   <- "Other"

df.T$Morality[grepl("moral", df.T$Shape, fixed=TRUE)]   <- "Good"
df.T$Morality[grepl("immoral", df.T$Shape, fixed=TRUE)] <- "Bad"

df.T$Identity[grepl("Self", df.T$Shape, fixed=TRUE)]    <- "Self"
df.T$Identity[grepl("Other", df.T$Shape, fixed=TRUE)]   <- "Other"

df.T$Task[df.T$Task == 'self']            <- 'Id'
df.T$Task[df.T$Task == 'moral']           <- 'Val'

# rename columns
colnames(df.L)[colnames(df.L)=="Sub"] <- "Subject"
colnames(df.T)[colnames(df.T)=="Sub"] <- "Subject"

# order the variables
df.L$Morality <- factor(df.L$Morality, levels = c("Good","Bad"))    
df.L$Identity <- factor(df.L$Identity, levels = c("Self","Other"))  
df.L$Match    <- factor(df.L$Match,    levels = c("match","mismatch"))

# make the variables in a specified order
df.T$Morality <- factor(df.T$Morality, levels=c("Good","Bad"))    
df.T$Identity <- factor(df.T$Identity, levels=c("Self","Other"))
```

```{r clean the data_e7,echo=FALSE,results='hide',warning=FALSE, message=FALSE}

# get the data file for hddm analysis
df.L.hddm <- subset(df.L, ACC ==1 | ACC == 0)  # remove non-response trials or pressing wrong key
df.T.hddm <- subset(df.T, ACC ==1 | ACC == 0)

df.L.hddm_m <- subset(df.L.hddm, Match == 'match') # select data
df.L.hddm_m <- df.L.hddm_m[,c('Subject','Morality','Identity','RT','ACC')] # select column
colnames(df.L.hddm_m) <- c("subj_idx","Morality",'id','rt','response')     # change column name

df.L.hddm_nm <- subset(df.L.hddm, Match == 'mismatch')
df.L.hddm_nm <- df.L.hddm_nm[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.L.hddm_nm) <- c("subj_idx","Morality",'id','rt','response')

df.T.hddm_val <- subset(df.T.hddm, Task == 'Val')
df.T.hddm_val <- df.T.hddm_val[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.T.hddm_val) <- c("subj_idx","Morality",'id','rt','response')
df.T.hddm_id <- subset(df.T.hddm, Task == 'Id')
df.T.hddm_id <- df.T.hddm_id[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.T.hddm_id) <- c("subj_idx","Morality",'id','rt','response')

setwd(ddmDir)
write.csv(df.L.hddm_m,'exp7_rep_match_hddm.csv',row.names = F)
write.csv(df.L.hddm_nm,'exp7_rep_mismatch_hddm.csv',row.names = F)
write.csv(df.T.hddm_val,'exp7_rep_categ_val_hddm.csv',row.names = F)
write.csv(df.T.hddm_id,'exp7_rep_categ_id_hddm.csv',row.names = F)
rm('df.L.hddm','df.T.hddm','df.L.hddm_m','df.L.hddm_nm','df.T.hddm_val','df.T.hddm_id')
setwd(curDir)

# clean data for trad. analysis
df.L1 <- df.L[,c("Subject", "Age", "Sex", "Match", "Morality", "Identity", "ACC","RT")]
df.T1 <- df.T[,c("Subject", "Age", "Sex", "Task", "Morality", "Identity", "ACC","RT")]
df.L1$RT <- df.L1$RT * 1000 # transfer from seconds to min seconds
df.T1$RT <- df.T1$RT * 1000 # careful about the scale of time when using HDDM or ex-Gaussian

# no response equal to wrong
df.L1$ACC[df.L1$ACC != 1] <- 0
df.T1$ACC[df.T1$ACC != 1] <- 0

# check the number of trials are correct
excldSub1_L <- c() # subject ID of those who failed matching task
excldSub1_T <- c() # subject ID of those who failed categorization task

df.L1.trials <-  plyr::ddply(df.L1,.(Subject, Match, Identity,Morality), summarise,
                    N = length(ACC),
                    countN = sum(ACC))
subjlistL <- unique(df.L1.trials$Subject)
for (ii in 1:length(subjlistL)){
  tmp <- subset(df.L1.trials,Subject == subjlistL[ii])
  if(sum(tmp$N != 75)){ # if the number of trials in matching task correct 
    excldSub1_L[ii] <- subjlistL[ii]
    print(paste0("ERROR: the trials number of subject ", subjlistL[ii] , 
                 " in matching task is not correct"))
}}

df.T1.trials <-  plyr::ddply(df.T1,.(Subject, Task, Identity,Morality), summarise,
                    N = length(ACC),
                    countN = sum(ACC))

subjlistT <- unique(df.T1.trials$Subject)
for (jj in 1:length(subjlistT)){
  tmp <- subset(df.T1.trials,Subject == subjlistT[jj])
  if(sum(tmp$N != 90)){ # if the number of trials in matching task correct 
    excldSub1_T[jj] <- subjlistT[jj]
    print(paste0("ERROR: the trials number of subject ", subjlistT[jj] , 
                 " in categ task is not correct"))
  }}

# get all the participants who were excluded:
excldSub1 <- c(excldSub1_L, excldSub1_T)
excldSub1 <- excldSub1[!duplicated(excldSub1)]

# df.L1 <- df.L1[!(df.L1$Subject %in% procFailureSub),]
# df.T1 <- df.T1[!(df.T1$Subject %in% procFailureSub),]

# calculate the overall accuracy for matching task
df.L1.acc.g <-  plyr::ddply(df.L1,.(Subject), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

# calculate the overall accuracy for categorziation task
df.T1.acc.g <-  plyr::ddply(df.T1,.(Subject), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

excldSub2_L <- df.L1.acc.g$Subject[df.L1.acc.g$ACC < 0.5]      # less 50% accuracy in matching task
excldSub2_T <- df.T1.acc.g$Subject[df.T1.acc.g$ACC < 0.5]      # less than 50% accuracy in categorization task
excldSub2   <- c(excldSub2_L, excldSub2_T)
excldSub2 <- excldSub2[!duplicated(excldSub2)]   # invalid because of low overall accuracy

excldSub   <- c(excldSub1, excldSub2)
excldSub <- excldSub[!duplicated(excldSub)]   # total excluded

valdSub <- setdiff(subjlistL, excldSub)

df.L1.valid <- df.L1[!(df.L1$Subject %in% excldSub),]   # exclude the invalid subjects
df.T1.valid <- df.T1[!(df.T1$Subject %in% excldSub),]
# df.T1.valid <- df.T1.valid[!(df.T1.valid$Subject %in% excld.sub.T),]

# check the number of participants are correct
length(valdSub) + length(excldSub) == length(unique(df.L1$Subject))

# excld.trials3 <- excld.trials[!(excld.trials$Subject %in% excld.sub),]
excld.trials.L <- df.L1.valid[df.L1.valid$RT <= 200 & df.L1.valid$ACC == 1,]
df.L1.V        <- df.L1.valid[!(df.L1.valid$RT <= 200 & df.L1.valid$ACC == 1),]  # valid trial data for match task
excld.trials.T <- df.T1.valid[df.T1.valid$RT <= 200 & df.T1.valid$ACC == 1,]
df.T1.V        <- df.T1.valid[!(df.T1.valid$RT <= 200 & df.T1.valid$ACC == 1),]  # valid trial data for categorization task
nrow(excld.trials.T) + nrow(df.T1.V) == nrow(df.T1.valid) # if true, everything is ok

## Basic information of the data ####
df.L1.T.basic <- df.L1[!duplicated(df.L1$Subject), 2:5]
numT.subj     <- nrow(df.L1.T.basic)
numT.female   <- sum(df.L1.T.basic$Sex == 'female');
numT.male     <- sum(df.L1.T.basic$Sex == 'male');
ageT.mean     <- round(mean(df.L1.T.basic$Age),2);
ageT.std      <- round(sd(df.L1.T.basic$Age),2);
num.excld.sub <- length(unique(excldSub))
# num.excld.sub.T <- length(unique(excld.sub.T))

# valide data for matching task
df.L1.V.basic <- df.L1.V[!duplicated(df.L1.V$Subject), 2:5]
numV.female   <- sum(df.L1.V.basic$Sex == 'female');
numV.male     <- sum(df.L1.V.basic$Sex == 'male');
ageV.mean     <- round(mean(df.L1.V.basic$Age),2);
ageV.std      <- round(sd(df.L1.V.basic$Age),2);
ratio.excld.trials.L <- nrow(excld.trials.L)/nrow(df.L1.valid)
ratio.excld.trials.T <- nrow(excld.trials.T)/nrow(df.T1.valid)

# valid data for categorization task
df.T1.V.basic <- df.T1.V[!duplicated(df.T1.V$Subject), 2:5]
numV.female.T <- sum(df.T1.V.basic$Sex == 'female');
numV.male.T <- sum(df.T1.V.basic$Sex == 'male');
ageV.mean.T <- round(mean(df.T1.V.basic$Age),2);
ageV.std.T <- round(sd(df.T1.V.basic$Age),2);
ratio.excld.trials.T <- nrow(excld.trials.T)/nrow(df.T1.valid)
```
## Participants
Total participants: `r numT.subj` (`r numT.female` female, age: `r ageT.mean` $\pm$ `r ageT.std`)

Excluded because of wrong trials number:

matching task: `r excldSub1_L` ;

categ. task: `r excldSub1_T`;

total: `r excldSub1`.

Excluded because the overall accuracy is lower than 50%: 
matching task: `r excldSub2_L` ;

categ. task: `r excldSub2_T`;

total: `r excldSub2`.

Remaining participant: `r length(valdSub)` (`r valdSub`)

```{r preproc_match,echo=FALSE,results='hide',warning=FALSE, message=FALSE}
### ACC ####
df.L1.V.acc  <-  ddply(df.L1.V,.(Subject, Match, Morality, Identity), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

df.L1.V.acc_w <- dcast(df.L1.V.acc, Subject ~ Match + Morality + Identity,value.var = "ACC")

# rename the column number
colnames(df.L1.V.acc_w)[2:9] <- paste("ACC", colnames(df.L1.V.acc_w[,2:9]), sep = "_")

# d prime #### 
df.L1.V$sdt <- NA
for (i in 1:nrow(df.L1.V)){
        if (df.L1.V$ACC[i] == 1 & df.L1.V$Match[i] == "match"){
                df.L1.V$sdt[i] <- "hit"
        } else if (df.L1.V$ACC[i] == 1 & df.L1.V$Match[i] == "mismatch"){
                df.L1.V$sdt[i] <- "CR"
        } else if (df.L1.V$ACC[i] == 0 & df.L1.V$Match[i] == "match"){
                df.L1.V$sdt[i] <- "miss"
        } else if (df.L1.V$ACC[i] == 0 & df.L1.V$Match[i] == "mismatch"){
                df.L1.V$sdt[i] <- "FA"
        }
}

# calculate the number of each for each condition
df.L1.V.SDT <-  ddply(df.L1.V,.(Subject,Age, Sex, Morality,Identity,sdt), summarise,
                     N = length(sdt))

# long format to wide
df.L1.V.SDT_w <- dcast(df.L1.V.SDT, Subject + Age + Sex+ Morality + Identity  ~ sdt,value.var = "N")
df.L1.V.SDT_w$miss[is.na(df.L1.V.SDT_w$miss)] <- 0
df.L1.V.SDT_w$FA[is.na(df.L1.V.SDT_w$FA)] <- 0
df.L1.V.SDT_w$hitR <- df.L1.V.SDT_w$hit/(df.L1.V.SDT_w$hit + df.L1.V.SDT_w$miss)
df.L1.V.SDT_w$faR <- df.L1.V.SDT_w$FA/(df.L1.V.SDT_w$FA + df.L1.V.SDT_w$CR)

# standardized way to deal with the extreme values
for (i in 1:nrow(df.L1.V.SDT_w)){
        if (df.L1.V.SDT_w$hitR[i] == 1){
                df.L1.V.SDT_w$hitR[i] <- 1 - 1/(2*(df.L1.V.SDT_w$hit[i] + df.L1.V.SDT_w$miss[i]))
        }
}

for (i in 1:nrow(df.L1.V.SDT_w)){
        if (df.L1.V.SDT_w$faR[i] == 0){
                df.L1.V.SDT_w$faR[i] <- 1/(2*(df.L1.V.SDT_w$FA[i] + df.L1.V.SDT_w$CR[i]))
        }
}

# calculate the d prime for each condition
df.L1.V.SDT_w$dprime <- mapply(dprime,df.L1.V.SDT_w$hitR,df.L1.V.SDT_w$faR)
df.L1.V.SDT_ww   <- dcast(df.L1.V.SDT_w, Subject + Sex + Age ~ Morality + Identity ,value.var = "dprime")

df.L1.V.SDT_l <- df.L1.V.SDT_w[,c(1:5,12)]

# rename the column number
colnames(df.L1.V.SDT_ww)[4:7] <- paste("d", colnames(df.L1.V.SDT_ww[,4:7]), sep = "_")

## doing the analysis for RT ####
df.L1.V.RT <- df.L1.V[df.L1.V$ACC == 1,]  # exclued inaccurate data
df.L1.V.RT.subj <- summarySEwithin(df.L1.V.RT,measurevar = 'RT', withinvar = c('Subject','Match','Morality','Identity'), idvar = 'Subject',na.rm = TRUE)
df.L1.V.RT.subj_w <- dcast(df.L1.V.RT.subj, Subject ~ Match + Morality + Identity ,value.var = "RT")

# rename the columns of RT data
colnames(df.L1.V.RT.subj_w)[2:9] <- paste("RT", colnames(df.L1.V.RT.subj_w[,2:9]), sep = "_")

## saving data ####
# merge the dprime and RT data and save
df.L1.V.sum_w <- merge(df.L1.V.acc_w,  df.L1.V.SDT_ww,by = "Subject")
df.L1.V.sum_w <- merge(df.L1.V.sum_w,df.L1.V.RT.subj_w,by = 'Subject')

# merge the RT and ACC data (long-format)
df.L1.V.sum_rt_acc_l <- merge(df.L1.V.acc,df.L1.V.RT.subj,by = c("Subject","Match","Morality",'Identity'))
df.L1.V.sum_rt_acc_l <- df.L1.V.sum_rt_acc_l[order(df.L1.V.sum_rt_acc_l$Subject),]

df.L1.V.sum_rt_acc_l <- df.L1.V.sum_rt_acc_l[,c("Subject","Match","Morality",'Identity',"N.x","countN","ACC","RT")]
colnames(df.L1.V.sum_rt_acc_l) <- c("Subject","Match","Morality",'Identity',"Ntrials","corrTrials","ACC","RT")

# order the columns
df.L1.V.sum_w <- df.L1.V.sum_w[,c(colnames(df.L1.V.sum_w)[c(1,10:11,2:9,12:23)])]

# write files
setwd(traDir)
write.csv(df.L1.V.sum_w,'exp7_rep_match_behav_wide.csv',row.names = F)
write.csv(df.L1.V.SDT_l,'exp7_rep_match__dprime_long.csv',row.names = F)
write.csv(df.L1.V.sum_rt_acc_l,'exp7_rep_match__rt_acc_long.csv',row.names = F)
setwd(curDir)

```

```{r preproc_categ,echo=FALSE,results='hide',warning=FALSE, message=FALSE}
### ACC ####
df.T1.V.acc  <-  ddply(df.T1.V,.(Subject, Task, Morality, Identity), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

df.T1.V.acc_w <- dcast(df.T1.V.acc, Subject ~ Task + Morality + Identity,value.var = "ACC")

# rename the column number
colnames(df.T1.V.acc_w)[2:9] <- paste("ACC", colnames(df.T1.V.acc_w[,2:9]), sep = "_")

## doing the analysis for RT ####
df.T1.V.RT <- df.T1.V[df.T1.V$ACC == 1,]  # exclued inaccurate data
df.T1.V.RT.subj <- summarySEwithin(df.T1.V.RT,measurevar = 'RT', withinvar = c('Subject','Task','Morality','Identity'), idvar = 'Subject',na.rm = TRUE)
df.T1.V.RT.subj_w <- dcast(df.T1.V.RT.subj, Subject ~ Task + Morality + Identity ,value.var = "RT")

# rename the columns of RT data
colnames(df.T1.V.RT.subj_w)[2:9] <- paste("RT", colnames(df.T1.V.RT.subj_w[,2:9]), sep = "_")

## saving data ####
# merge the dprime and RT data and save
df.T1.V.sum_w <- merge(df.T1.V.acc_w,  df.T1.V.RT.subj_w,by = "Subject")

# merge the RT and ACC data (long-format)
df.T1.V.sum_rt_acc_l <- merge(df.T1.V.acc,df.T1.V.RT.subj,by = c("Subject","Task","Morality",'Identity'))
df.T1.V.sum_rt_acc_l <- df.T1.V.sum_rt_acc_l[order(df.T1.V.sum_rt_acc_l$Subject),]

df.T1.V.sum_rt_acc_l <- df.T1.V.sum_rt_acc_l[,c("Subject","Task","Morality",'Identity',"N.x","countN","ACC","RT")]
colnames(df.T1.V.sum_rt_acc_l) <- c("Subject","Task","Morality",'Identity',"Ntrials","corrTrials","ACC","RT")

# order the columns
#df.T1.V.sum_w <- df.T1.V.sum_w[,c(colnames(df.T1.V.sum_w)[c(1,10:11,2:9,12:23)])]

# write files to an upper-lelel folder
setwd(traDir)
write.csv(df.T1.V.sum_w,'exp7_rep_categ_behav_wide.csv',row.names = F)
write.csv(df.T1.V.sum_rt_acc_l,'exp7_rep_categ__rt_acc_long.csv',row.names = F)
setwd(curDir)

```

```{r plot_ind_data,echo=FALSE,warning=FALSE, message=FALSE}
df.L1.V.sum_rt_acc_l_p <- subset(df.L1.V.sum_rt_acc_l,Match == 'match')
df.L1.V.sum_rt_acc_l_p$cond <- paste(df.L1.V.sum_rt_acc_l_p$Morality,df.L1.V.sum_rt_acc_l_p$Identity,sep = '_')
df.L1.V.sum_rt_acc_l_p <- subset(df.L1.V.sum_rt_acc_l_p,Morality == 'Good')
ggplot(data=df.L1.V.sum_rt_acc_l_p, aes(x=cond, y=RT, group=Subject, colour=Subject)) +
    geom_line() +
    geom_point() +
    apatheme

# plot the d prime for good other and good self
# concatenate the conditions
df.L1.V.SDT_l$cond <- paste(df.L1.V.SDT_l$Morality,df.L1.V.SDT_l$Identity, sep = '_')
# select only the moral positive coniditions
df.L1.V.SDT_l_p <- df.L1.V.SDT_l[df.L1.V.SDT_l$Morality == 'Good',]
# plot 
ggplot(data=df.L1.V.SDT_l_p, aes(x=cond, y=dprime, group=Subject, colour=Subject)) +
    geom_line() +
    geom_point() +
    apatheme

#ggstatsplot::ggbetweenstats(
#  data = df.L1.V.sum_rt_acc_l_p,
#  x = c(Morality,Identity),
#  y = RT,
#  notch = TRUE,                                   # show notched box plot
#  mean.plotting = TRUE,                           # whether mean for each group id to be displayed 
#  #type = "parametric",                            # which type of test is to be run
#  outlier.tagging = FALSE,                         # whether outliers need to be tagged
#  #outlier.label = RT,                    # variable to be used for the outlier tag
#  xlab = "conditions",                       # label for the x-axis variable
#  ylab = "Reaction time (ms)",               # label for the y-axis variable
#  title = "Dataset: Iris flower data set",        # title text for the plot
#  caption = expression(                           # caption text for the plot 
#    paste(italic("Note"), ": this is a demo")
#  ),
#  messages = FALSE
#) +                                               # further modification outside of ggstatsplot
#  ggplot2::coord_cartesian(ylim = c(500, 800)) + 
#  ggplot2::scale_y_continuous(breaks = seq(500, 800, by = 50)) 
```