### 2018-06-26 Replication/Results/data_analysis/preproc/Analysis_exp7_prepro.rmd
add separate script to save raw data

---
title: "Preprocessing_&_Plot_for_Moral_Categ_Replication"
author: "hcp"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
---
<style type="text/css">

body{ /* Normal  */
   font-family: Times     
   font-size: 12px;
}
td {  /* Table  */
   font-size: 8px;
}
h1 { /* Header 1 */
 font-size: 28px;
}
h2 { /* Header 2 */
 font-size: 22px;
}
h3 { /* Header 3 */
 font-size: 18px;
 color: DarkBlue;
}
code.r{ /* Code block */
  font-size: 10px;
}
pre { /* Code block */
  font-size: 10px
}
</style>

This script is used for quality control and preprocessing of the replication data.

```{r Initializing, include=FALSE,,echo=FALSE,results='hide',warning=FALSE, message=FALSE}
source('Initial.r')

# set the directories
curDir <- getwd()   # directory for preprocessing
rootDir <- gsub('.{7}$', '', curDir)
rawDir <- paste(curDir,'/data/',sep = '')
traDir <- paste(rootDir,'tradAnal',sep = '')
ddmDir <- paste(rootDir,'hddmMod',sep = '')
```

```{r loadingData_e7,,echo=FALSE,results='hide',warning=FALSE, message=FALSE}
# read these files and combine them into one file, get the raw data for matching task
df.M <- read.csv('MS_rep_matchingTask_raw.csv', header=TRUE, sep=",",stringsAsFactors=F)
df.C <- read.csv('MS_rep_categTask_raw.csv', header=TRUE, sep=",",stringsAsFactors=F)
```

```{r clean the data_e7,echo=FALSE,results='hide',warning=FALSE, message=FALSE}

# clean data for trad. analysis
df.M1 <- df.M
df.C1 <- df.C
df.M1$RT <- df.M1$RT * 1000 # transfer from seconds to min seconds
df.C1$RT <- df.C1$RT * 1000 # careful about the scale of time when using HDDM or ex-Gaussian

# no response equal to wrong
df.M1$ACC[df.M1$ACC != 1] <- 0
df.C1$ACC[df.C1$ACC != 1] <- 0

# check the number of trials are correct
excldSub1_M  <- c()     # subject ID of those who failed matching task
excldSub1_C  <- c(7338) # subject 7338 excluded because Accuracy of one condition is zero

df.M1.trials <-  plyr::ddply(df.M1,.(Subject, Match, Identity,Morality), summarise,
                    N = length(ACC),
                    countN = sum(ACC))
subjlistM <- unique(df.M1.trials$Subject)
excldSub2_M <- c()
for (ii in 1:length(subjlistM)){
  tmp <- subset(df.M1.trials,Subject == subjlistM[ii])
  if(sum(tmp$N != 75)){ # if the number of trials in matching task correct 
    excldSub2_M[ii] <- subjlistM[ii]
    print(paste0("ERROR: the trials number of subject ", subjlistM[ii] , 
                 " in matching task is not correct"))
}}

df.C1.trials <-  plyr::ddply(df.C1,.(Subject, Task, Identity,Morality), summarise,
                    N = length(ACC),
                    countN = sum(ACC))

subjlistC <- unique(df.C1.trials$Subject)
excldSub2_C <- c()
for (jj in 1:length(subjlistC)){
  tmp <- subset(df.C1.trials,Subject == subjlistC[jj])
  if(sum(tmp$N != 90)){ # if the number of trials in matching task correct 
    excldSub2_C[jj] <- subjlistC[jj]
    print(paste0("ERROR: the trials number of subject ", subjlistC[jj] , 
                 " in categ task is not correct"))
  }}

# df.M1 <- df.M1[!(df.M1$Subject %in% procFailureSub),]
# df.C1 <- df.C1[!(df.C1$Subject %in% procFailureSub),]

# calculate the overall accuracy for matching task
df.M1.acc.g <-  plyr::ddply(df.M1,.(Subject), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

# calculate the overall accuracy for categorziation task
df.C1.acc.g <-  plyr::ddply(df.C1,.(Subject), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

excldSub3_M <- df.M1.acc.g$Subject[df.M1.acc.g$ACC < 0.5]      # less 50% accuracy in matching task
excldSub3_C <- df.C1.acc.g$Subject[df.C1.acc.g$ACC < 0.5]      # less than 50% accuracy in categorization task

excldSub_M   <- c(excldSub1_M, excldSub2_M,excldSub3_M)
excldSub_M <- excldSub_M[!duplicated(excldSub_M)]   # total excluded for matching task

excldSub_C   <- c(excldSub1_C, excldSub2_C,excldSub3_C)
excldSub_C <- excldSub_C[!duplicated(excldSub_C)]   # total excluded for matching task

valdSub_M <- setdiff(subjlistM, excldSub_M)
valdSub_C <- setdiff(subjlistC, excldSub_C)

# select valid data for further analysis

df.M1.valid <- df.M1[!(df.M1$Subject %in% excldSub_M),]   # exclude the invalid subjects
df.C1.valid <- df.C1[!(df.C1$Subject %in% excldSub_C),]
# df.C1.valid <- df.C1.valid[!(df.C1.valid$Subject %in% excld.sub.T),]

# check the number of participants are correct
length(valdSub_M) + length(excldSub_M) == length(unique(df.M1$Subject))
length(valdSub_C) + length(excldSub_C) == length(unique(df.C1$Subject))

# excluded correct trials with < 200ms RT
excld.trials.M <- df.M1.valid[df.M1.valid$RT <= 200 & df.M1.valid$ACC == 1,]
df.M1.V        <- df.M1.valid[!(df.M1.valid$RT <= 200 & df.M1.valid$ACC == 1),]  # valid trial data for match task
excld.trials.C <- df.C1.valid[df.C1.valid$RT <= 200 & df.C1.valid$ACC == 1,]
df.C1.V        <- df.C1.valid[!(df.C1.valid$RT <= 200 & df.C1.valid$ACC == 1),]  # valid trial data for categorization task
nrow(excld.trials.C) + nrow(df.C1.V) == nrow(df.C1.valid) # if true, everything is ok

## Basic information of the data ####
df.M1.T.basic <- df.M1[!duplicated(df.M1$Subject), 2:5]
numT.subj     <- nrow(df.M1.T.basic)
numT.female   <- sum(df.M1.T.basic$Sex == 'female');
numT.male     <- sum(df.M1.T.basic$Sex == 'male');
ageT.mean     <- round(mean(df.M1.T.basic$Age),2);
ageT.std      <- round(sd(df.M1.T.basic$Age),2);
# num.excld.sub <- length(unique(excldSub))
# num.excld.sub.T <- length(unique(excld.sub.T))

# valide data for matching task
df.M1.V.basic <- df.M1.V[!duplicated(df.M1.V$Subject), 2:5]
numV.female   <- sum(df.M1.V.basic$Sex == 'female');
numV.male     <- sum(df.M1.V.basic$Sex == 'male');
ageV.mean     <- round(mean(df.M1.V.basic$Age),2);
ageV.std      <- round(sd(df.M1.V.basic$Age),2);
ratio.excld.trials.M <- nrow(excld.trials.M)/nrow(df.M1.valid)
num.excld.sub_M <- length(unique(valdSub_M))

# valid data for categorization task
df.C1.V.basic <- df.C1.V[!duplicated(df.C1.V$Subject), 2:5]
numV.female.C <- sum(df.C1.V.basic$Sex == 'female');
numV.male.C <- sum(df.C1.V.basic$Sex == 'male');
ageV.mean.C <- round(mean(df.C1.V.basic$Age),2);
ageV.std.C <- round(sd(df.C1.V.basic$Age),2);
ratio.excld.trials.C <- nrow(excld.trials.C)/nrow(df.C1.valid)
num.excld.sub_C <- length(unique(valdSub_C))

# get the data file for hddm analysis

df.M.ddm.v <- df.M[!(df.M$Subject %in% excldSub_M),]   # exclude the invalid subjects
df.C.ddm.v <- df.C[!(df.C$Subject %in% excldSub_C),]

df.M.hddm <- subset(df.M.ddm.v, ACC ==1 | ACC == 0)  # remove non-response trials or pressing wrong key
df.C.hddm <- subset(df.C.ddm.v, ACC ==1 | ACC == 0)

df.M.hddm_m <- subset(df.M.hddm, Match == 'match') # select data
df.M.hddm_m <- df.M.hddm_m[,c('Subject','Morality','Identity','RT','ACC')] # select column
colnames(df.M.hddm_m) <- c("subj_idx","val",'id','rt','response')     # change column name

df.M.hddm_nm <- subset(df.M.hddm, Match == 'mismatch')
df.M.hddm_nm <- df.M.hddm_nm[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.M.hddm_nm) <- c("subj_idx","val",'id','rt','response')

df.C.hddm_val <- subset(df.C.hddm, Task == 'Val')
df.C.hddm_val <- df.C.hddm_val[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.C.hddm_val) <- c("subj_idx","val",'id','rt','response')

df.C.hddm_id <- subset(df.C.hddm, Task == 'Id')
df.C.hddm_id <- df.C.hddm_id[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.C.hddm_id) <- c("subj_idx","val",'id','rt','response')

setwd(ddmDir)
write.csv(df.M.hddm_m,'exp7_rep_match_hddm.csv',row.names = F)
write.csv(df.M.hddm_nm,'exp7_rep_mismatch_hddm.csv',row.names = F)
write.csv(df.C.hddm_val,'exp7_rep_categ_val_hddm.csv',row.names = F)
write.csv(df.C.hddm_id,'exp7_rep_categ_id_hddm.csv',row.names = F)
rm('df.M.hddm','df.C.hddm','df.M.hddm_m','df.M.hddm_nm','df.C.hddm_val','df.C.hddm_id')
setwd(curDir)

```
## Participants
Total participants: `r numT.subj` (`r numT.female` female, age: `r ageT.mean` $\pm$ `r ageT.std`)

Excluded because of wrong trials number:

matching task: `r excldSub1_M` ;

categ. task: `r excldSub1_C`.

Excluded because the overall accuracy is lower than 50%: 
matching task: `r excldSub2_M` ;

categ. task: `r excldSub2_C`.

Remaining participant:

matching task`r length(valdSub_M)`; categorization task: (`r length(valdSub_C)`)

```{r preproc_match,echo=FALSE,results='hide',warning=FALSE, message=FALSE}
### ACC ####
df.M1.V.acc  <-  plyr::ddply(df.M1.V,.(Subject, Match, Morality, Identity), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

df.M1.V.acc_w <- dcast(df.M1.V.acc, Subject ~ Match + Morality + Identity,value.var = "ACC")

# rename the column number
colnames(df.M1.V.acc_w)[2:9] <- paste("ACC", colnames(df.M1.V.acc_w[,2:9]), sep = "_")

# d prime #### 
df.M1.V$sdt <- NA
for (i in 1:nrow(df.M1.V)){
        if (df.M1.V$ACC[i] == 1 & df.M1.V$Match[i] == "match"){
                df.M1.V$sdt[i] <- "hit"
        } else if (df.M1.V$ACC[i] == 1 & df.M1.V$Match[i] == "mismatch"){
                df.M1.V$sdt[i] <- "CR"
        } else if (df.M1.V$ACC[i] == 0 & df.M1.V$Match[i] == "match"){
                df.M1.V$sdt[i] <- "miss"
        } else if (df.M1.V$ACC[i] == 0 & df.M1.V$Match[i] == "mismatch"){
                df.M1.V$sdt[i] <- "FA"
        }
}

# calculate the number of each for each condition
df.M1.V.SDT <-  plyr::ddply(df.M1.V,.(Subject,Age, Sex, Morality,Identity,sdt), summarise,
                     N = length(sdt))

# long format to wide
df.M1.V.SDT_w <- dcast(df.M1.V.SDT, Subject + Age + Sex+ Morality + Identity  ~ sdt,value.var = "N")
df.M1.V.SDT_w$miss[is.na(df.M1.V.SDT_w$miss)] <- 0
df.M1.V.SDT_w$FA[is.na(df.M1.V.SDT_w$FA)] <- 0
df.M1.V.SDT_w$hitR <- df.M1.V.SDT_w$hit/(df.M1.V.SDT_w$hit + df.M1.V.SDT_w$miss)
df.M1.V.SDT_w$faR <- df.M1.V.SDT_w$FA/(df.M1.V.SDT_w$FA + df.M1.V.SDT_w$CR)

# standardized way to deal with the extreme values
for (i in 1:nrow(df.M1.V.SDT_w)){
        if (df.M1.V.SDT_w$hitR[i] == 1){
                df.M1.V.SDT_w$hitR[i] <- 1 - 1/(2*(df.M1.V.SDT_w$hit[i] + df.M1.V.SDT_w$miss[i]))
        }
}

for (i in 1:nrow(df.M1.V.SDT_w)){
        if (df.M1.V.SDT_w$faR[i] == 0){
                df.M1.V.SDT_w$faR[i] <- 1/(2*(df.M1.V.SDT_w$FA[i] + df.M1.V.SDT_w$CR[i]))
        }
}

# calculate the d prime for each condition
df.M1.V.SDT_w$dprime <- mapply(dprime,df.M1.V.SDT_w$hitR,df.M1.V.SDT_w$faR)
df.M1.V.SDT_ww   <- dcast(df.M1.V.SDT_w, Subject + Sex + Age ~ Morality + Identity ,value.var = "dprime")

df.M1.V.SDT_l <- df.M1.V.SDT_w[,c(1:5,12)]

# rename the column number
colnames(df.M1.V.SDT_ww)[4:7] <- paste("d", colnames(df.M1.V.SDT_ww[,4:7]), sep = "_")

## doing the analysis for RT ####
df.M1.V.RT <- df.M1.V[df.M1.V$ACC == 1,]  # exclued inaccurate data
df.M1.V.RT.subj <- summarySEwithin(df.M1.V.RT,measurevar = 'RT', withinvar = c('Subject','Match','Morality','Identity'), idvar = 'Subject',na.rm = TRUE)
df.M1.V.RT.subj_w <- dcast(df.M1.V.RT.subj, Subject ~ Match + Morality + Identity ,value.var = "RT")

# rename the columns of RT data
colnames(df.M1.V.RT.subj_w)[2:9] <- paste("RT", colnames(df.M1.V.RT.subj_w[,2:9]), sep = "_")

## saving data ####
# merge the dprime and RT data and save
df.M1.V.sum_w <- merge(df.M1.V.acc_w,  df.M1.V.SDT_ww,by = "Subject")
df.M1.V.sum_w <- merge(df.M1.V.sum_w,df.M1.V.RT.subj_w,by = 'Subject')

# merge the RT and ACC data (long-format)
df.M1.V.sum_rt_acc_l <- merge(df.M1.V.acc,df.M1.V.RT.subj,by = c("Subject","Match","Morality",'Identity'))
df.M1.V.sum_rt_acc_l <- df.M1.V.sum_rt_acc_l[order(df.M1.V.sum_rt_acc_l$Subject),]

df.M1.V.sum_rt_acc_l <- df.M1.V.sum_rt_acc_l[,c("Subject","Match","Morality",'Identity',"N.x","countN","ACC","RT")]
colnames(df.M1.V.sum_rt_acc_l) <- c("Subject","Match","Morality",'Identity',"Ntrials","corrTrials","ACC","RT")

# order the columns
df.M1.V.sum_w <- df.M1.V.sum_w[,c(colnames(df.M1.V.sum_w)[c(1,10:11,2:9,12:23)])]

# calculate the effect of self-ref and valence
df.M1.v.sum_eff_w <- df.M1.V.sum_w[,1:3]
df.M1.v.sum_eff_w$d_Good_Self_Other <- df.M1.V.sum_w$d_Good_Self - df.M1.V.sum_w$d_Good_Other
df.M1.v.sum_eff_w$d_Self_Good_Bad   <- df.M1.V.sum_w$d_Good_Self - df.M1.V.sum_w$d_Bad_Self

df.M1.v.sum_eff_w$RT_Good_Self_Other <- df.M1.V.sum_w$RT_match_Good_Other -  df.M1.V.sum_w$RT_match_Good_Self
df.M1.v.sum_eff_w$RT_Self_Good_Bad   <- df.M1.V.sum_w$RT_match_Bad_Self -  df.M1.V.sum_w$RT_match_Good_Self

# write files
setwd(traDir)
write.csv(df.M1.V.sum_w,'exp7_rep_match_behav_wide.csv',row.names = F)
write.csv(df.M1.V.SDT_l,'exp7_rep_match__dprime_long.csv',row.names = F)
write.csv(df.M1.V.sum_rt_acc_l,'exp7_rep_match__rt_acc_long.csv',row.names = F)
setwd(curDir)

```

```{r preproc_categ,echo=FALSE,results='hide',warning=FALSE, message=FALSE}
### ACC ####
df.C1.V.acc  <-  plyr::ddply(df.C1.V,.(Subject, Task, Morality, Identity), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

df.C1.V.acc_w <- dcast(df.C1.V.acc, Subject ~ Task + Morality + Identity,value.var = "ACC")

# rename the column number
colnames(df.C1.V.acc_w)[2:9] <- paste("ACC", colnames(df.C1.V.acc_w[,2:9]), sep = "_")

# combing data from diff task for analyzing the interaction btw val and id
df.C1.V.acc_noTask  <-  plyr::ddply(df.C1.V,.(Subject, Morality, Identity), summarise,
                    N = length(ACC),
                    countN = sum(ACC),
                    ACC = sum(ACC)/length(ACC))

df.C1.V.acc_noTask_w <- dcast(df.C1.V.acc_noTask, Subject ~ Morality + Identity,value.var = "ACC")

# rename the column number
colnames(df.C1.V.acc_noTask_w)[2:5] <- paste("ACC", colnames(df.C1.V.acc_noTask_w[,2:5]), sep = "_")

## doing the analysis for RT ####
df.C1.V.RT <- df.C1.V[df.C1.V$ACC == 1,]  # exclued inaccurate data
df.C1.V.RT.subj <- summarySEwithin(df.C1.V.RT,measurevar = 'RT', withinvar = c('Subject','Task','Morality','Identity'), idvar = 'Subject',na.rm = TRUE)
df.C1.V.RT.subj_w <- dcast(df.C1.V.RT.subj, Subject ~ Task + Morality + Identity ,value.var = "RT")

# rename the columns of RT data
colnames(df.C1.V.RT.subj_w)[2:9] <- paste("RT", colnames(df.C1.V.RT.subj_w[,2:9]), sep = "_")

# combining data form different task for analyszing interaction of val and id
df.C1.V.RT.subj_noTask <- summarySEwithin(df.C1.V.RT,measurevar = 'RT', withinvar = c('Subject','Morality','Identity'), idvar = 'Subject',na.rm = TRUE)
df.C1.V.RT.subj_noTask_w <- dcast(df.C1.V.RT.subj_noTask, Subject ~ Morality + Identity ,value.var = "RT")

# rename the columns of RT data
colnames(df.C1.V.RT.subj_noTask_w)[2:5] <- paste("RT", colnames(df.C1.V.RT.subj_noTask_w[,2:5]), sep = "_")

## saving data ####
# merge the accuracy and RT data and save
df.C1.V.sum_w <- merge(df.C1.V.acc_w,  df.C1.V.RT.subj_w,by = "Subject")
df.C1.V.sum_noTask_w <- merge(df.C1.V.acc_noTask_w,  df.C1.V.RT.subj_noTask_w,by = "Subject")

## calculated effects ####
# self-reference effect
df.C1.V.sum_w$RT_Val_Good_selfEffect <- df.C1.V.sum_w$RT_Val_Good_Other - df.C1.V.sum_w$RT_Val_Good_Self
df.C1.V.sum_w$RT_Val_Bad_selfEffect  <- df.C1.V.sum_w$RT_Val_Bad_Other - df.C1.V.sum_w$RT_Val_Bad_Self
df.C1.V.sum_w$RT_Id_Good_selfEffect  <- df.C1.V.sum_w$RT_Id_Good_Other - df.C1.V.sum_w$RT_Id_Good_Self
df.C1.V.sum_w$RT_Id_Bad_selfEffect   <- df.C1.V.sum_w$RT_Id_Bad_Other - df.C1.V.sum_w$RT_Id_Bad_Self
# valence effect
df.C1.V.sum_w$RT_Val_Self_valEffect  <- df.C1.V.sum_w$RT_Val_Bad_Self - df.C1.V.sum_w$RT_Val_Good_Self
df.C1.V.sum_w$RT_Val_Other_valEffect <- df.C1.V.sum_w$RT_Val_Bad_Other - df.C1.V.sum_w$RT_Val_Good_Other
df.C1.V.sum_w$RT_Id_Self_valEffect   <- df.C1.V.sum_w$RT_Id_Bad_Self - df.C1.V.sum_w$RT_Id_Good_Self
df.C1.V.sum_w$RT_Id_Other_valEffect  <- df.C1.V.sum_w$RT_Id_Bad_Other - df.C1.V.sum_w$RT_Id_Good_Other
# task effect
df.C1.V.sum_w$RT_Good_Self_taskEffect  <- df.C1.V.sum_w$RT_Val_Good_Self - df.C1.V.sum_w$RT_Id_Good_Self
df.C1.V.sum_w$RT_Bad_Self_taskEffect   <- df.C1.V.sum_w$RT_Val_Bad_Self - df.C1.V.sum_w$RT_Id_Bad_Self
df.C1.V.sum_w$RT_Good_Other_taskEffect <- df.C1.V.sum_w$RT_Val_Good_Self - df.C1.V.sum_w$RT_Id_Good_Other
df.C1.V.sum_w$RT_Bad_Other_taskEffect  <- df.C1.V.sum_w$RT_Val_Bad_Other - df.C1.V.sum_w$RT_Id_Bad_Other

# calculate the effect of self-ref and valence
df.C1.v.sum_eff_w <- data.frame(df.C1.V.sum_w[,c('Subject')])
colnames(df.C1.v.sum_eff_w) <- 'Subject'
df.C1.v.sum_eff_w$RT_Val_Good_selfEffect <- df.C1.V.sum_w$RT_Val_Good_Other - df.C1.V.sum_w$RT_Val_Good_Self
df.C1.v.sum_eff_w$RT_Val_Self_valEffect  <- df.C1.V.sum_w$RT_Val_Bad_Self - df.C1.V.sum_w$RT_Val_Good_Self
df.C1.v.sum_eff_w$RT_Id_Good_selfEffect <- df.C1.V.sum_w$RT_Id_Good_Other - df.C1.V.sum_w$RT_Id_Good_Self
df.C1.v.sum_eff_w$RT_Id_Self_valEffect  <- df.C1.V.sum_w$RT_Id_Bad_Self - df.C1.V.sum_w$RT_Id_Good_Self

df.C1.v.sum_eff_w$ACC_Val_Good_selfEffect <- df.C1.V.sum_w$ACC_Val_Good_Other - df.C1.V.sum_w$ACC_Val_Good_Self
df.C1.v.sum_eff_w$ACC_Val_Self_valEffect  <- df.C1.V.sum_w$ACC_Val_Bad_Self - df.C1.V.sum_w$ACC_Val_Good_Self
df.C1.v.sum_eff_w$ACC_Id_Good_selfEffect <- df.C1.V.sum_w$ACC_Id_Good_Other - df.C1.V.sum_w$ACC_Id_Good_Self
df.C1.v.sum_eff_w$ACC_Id_Self_valEffect  <- df.C1.V.sum_w$ACC_Id_Bad_Self - df.C1.V.sum_w$ACC_Id_Good_Self

# merge the effect file
df.v.sum_eff_all_w <- merge(df.M1.v.sum_eff_w,df.C1.v.sum_eff_w,by="Subject")

# merge the RT and ACC data (long-format) ####
df.C1.V.sum_rt_acc_l <- merge(df.C1.V.acc,df.C1.V.RT.subj,by = c("Subject","Task","Morality",'Identity'))
df.C1.V.sum_rt_acc_l <- df.C1.V.sum_rt_acc_l[order(df.C1.V.sum_rt_acc_l$Subject),]

df.C1.V.sum_rt_acc_l <- df.C1.V.sum_rt_acc_l[,c("Subject","Task","Morality",'Identity',"N.x","countN","ACC","RT")]
colnames(df.C1.V.sum_rt_acc_l) <- c("Subject","Task","Morality",'Identity',"Ntrials","corrTrials","ACC","RT")

# order the columns
#df.C1.V.sum_w <- df.C1.V.sum_w[,c(colnames(df.C1.V.sum_w)[c(1,10:11,2:9,12:23)])]

# write files to an upper-lelel folder
setwd(traDir)
write.csv(df.C1.V.sum_w,'exp7_rep_categ_behav_wide.csv',row.names = F)
write.csv(df.C1.V.sum_noTask_w,'exp7_rep_categ_behav_noTask_wide.csv',row.names = F)
write.csv(df.C1.V.sum_rt_acc_l,'exp7_rep_categ__rt_acc_long.csv',row.names = F)
write.csv(df.v.sum_eff_all_w,'exp7_cross_taskeffect_wide.csv',row.names = F)
setwd(curDir)
```

```{r plot_match_RT_data,echo=FALSE,warning=FALSE, message=FALSE}
RTdata1 <- df.M1.V.sum_rt_acc_l %>%
  select(Subject,Match,Morality,Identity,RT) %>% 
  filter(Match == "match" & Morality == "Good")
RTdata1$Identity <- factor(RTdata1$Identity,levels = c("Self","Other"))
p1 <- ggplot(data = RTdata1, aes(y = RT, x = Identity,fill = Identity)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0)) +
  geom_point(aes(y = RT,color = Identity), position = position_jitter(width = .15), size = .5) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  guides(fill = FALSE) +guides(color = FALSE)+
  theme_bw() +
  raincloud_theme+scale_y_continuous(limits = c(300,900))+labs(x = "Good",y = "Reaction times(ms)")
RTdata2 <- df.M1.V.sum_rt_acc_l %>%
  select(Subject,Match,Morality,Identity,RT) %>% 
  filter(Match == "match" & Morality == "Bad")
RTdata2$Identity <- factor(RTdata2$Identity,levels = c("Self","Other"))
p2 <- ggplot(data = RTdata2, aes(y = RT, x = Identity,fill = Identity)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0)) +
  geom_point(aes(y = RT,color = Identity), position = position_jitter(width = .15), size = .5) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  guides(fill = FALSE) +guides(color = FALSE)+
  theme_bw() +
  raincloud_theme+scale_y_continuous(limits = c(300,900),breaks = NULL)+labs(x = "Bad",y = "")

setwd(traDir)
tiff("p_rt_match.tiff", width = 9, height = 6, units = 'in', res = 300)
p_rt_match <- multiplot(p1,p2,cols = 2)
dev.off()
setwd(curDir)
multiplot(p1,p2,cols = 2)
```

```{r plot_match_dprime_data,echo=FALSE,warning=FALSE, message=FALSE}
Ddata1 <- df.M1.V.SDT_l %>%
  select(Subject,Morality,Identity,dprime) %>% 
  filter(Morality == "Good")
Ddata1$Identity <- factor(Ddata1$Identity,levels = c("Self","Other"))
p1 <- ggplot(data = Ddata1, aes(y = dprime, x = Identity,fill = Identity)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0)) +
  geom_point(aes(y = dprime,color = Identity), position = position_jitter(width = .15), size = .5) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  guides(fill = FALSE) +guides(color = FALSE)+
  theme_bw() +
  raincloud_theme+scale_y_continuous(limits = c(0,4))+labs(x = "Good",y = "dprime")
Ddata2 <- df.M1.V.SDT_l %>%
  select(Subject,Morality,Identity,dprime) %>% 
  filter(Morality == "Bad")
Ddata2$Identity <- factor(RTdata2$Identity,levels = c("Self","Other"))
p2 <- ggplot(data = Ddata2, aes(y = dprime, x = Identity,fill = Identity)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0)) +
  geom_point(aes(y = dprime,color = Identity), position = position_jitter(width = .15), size = .5) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  guides(fill = FALSE) +guides(color = FALSE)+
  theme_bw() +
  raincloud_theme+scale_y_continuous(limits = c(0,4),breaks = NULL)+labs(x = "Bad",y = "")
p_dprime_match <- multiplot(p1,p2,cols = 2)
# ggsave("p_rt_match.pdf", p_rt_match, scale = 1,height = 6, width = 8, dpi = 300, family = "Times")
setwd(traDir)
tiff("p_dprime_match.tiff", width = 9, height = 6, units = 'in', res = 300)
p_dprime_match <- multiplot(p1,p2,cols = 2)
dev.off()
setwd(curDir)
#multiplot(p1,p2,cols = 2)
```

```{r plot_categ_ACC_val_data,echo=FALSE,warning=FALSE, message=FALSE}
ACCdata1 <- df.C1.V.sum_rt_acc_l %>%
  select(Subject,Task,Morality,Identity,ACC) %>% 
  filter(Task == "Val"& Morality == 'Good')
ACCdata1$Identity <- factor(ACCdata1$Identity,levels = c("Self","Other"))
p1 <- ggplot(data = ACCdata1, aes(y = ACC, x = Identity,fill = Identity)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0)) +
  geom_point(aes(y = ACC,color = Identity), position = position_jitter(width = .15), size = .5) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  guides(fill = FALSE) +guides(color = FALSE)+
  theme_bw() +
  raincloud_theme+scale_y_continuous(limits = c(0,1))+labs(x = "Good",y = "Accuracy")

ACCdata2 <- df.C1.V.sum_rt_acc_l %>%
  select(Subject,Task,Morality,Identity,ACC) %>% 
  filter(Task == "Val"& Morality == 'Bad')
ACCdata2$Identity <- factor(ACCdata2$Identity,levels = c("Self","Other"))
p2 <- ggplot(data = ACCdata2, aes(y = ACC, x = Identity,fill = Identity)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0)) +
  geom_point(aes(y = ACC,color = Identity), position = position_jitter(width = .15), size = .5) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  guides(fill = FALSE) +guides(color = FALSE)+
  theme_bw() +
  raincloud_theme+scale_y_continuous(limits = c(0,1),breaks = NULL)+labs(x = "Bad",y = "")
setwd(traDir)
tiff("p_rt_match.tiff", width = 9, height = 6, units = 'in', res = 300)
p_rt_match <- multiplot(p1,p2,cols = 2)
dev.off()
setwd(curDir)
multiplot(p1,p2,cols = 2)
```

```{r plot_categ_ACC_ID_data,echo=FALSE,warning=FALSE, message=FALSE}
ACCdata1 <- df.C1.V.sum_rt_acc_l %>%
  select(Subject,Task,Morality,Identity,ACC) %>% 
  filter(Task == "Id"& Morality == 'Good')
ACCdata1$Identity <- factor(ACCdata1$Identity,levels = c("Self","Other"))
p1 <- ggplot(data = ACCdata1, aes(y = ACC, x = Identity,fill = Identity)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0)) +
  geom_point(aes(y = ACC,color = Identity), position = position_jitter(width = .15), size = .5) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  guides(fill = FALSE) +guides(color = FALSE)+
  theme_bw() +
  raincloud_theme+scale_y_continuous(limits = c(0,1))+labs(x = "Good",y = "Accuracy")
ACCdata2 <- df.C1.V.sum_rt_acc_l %>%
  select(Subject,Task,Morality,Identity,ACC) %>% 
  filter(Task == "Id"& Morality == 'Bad')
ACCdata2$Identity <- factor(ACCdata2$Identity,levels = c("Self","Other"))
p2 <- ggplot(data = ACCdata2, aes(y = ACC, x = Identity,fill = Identity)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0)) +
  geom_point(aes(y = ACC,color = Identity), position = position_jitter(width = .15), size = .5) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  guides(fill = FALSE) +guides(color = FALSE)+
  theme_bw() +
  raincloud_theme+scale_y_continuous(limits = c(0,1),breaks = NULL)+labs(x = "Bad",y = "")
setwd(traDir)
tiff("p_rt_match.tiff", width = 9, height = 6, units = 'in', res = 300)
p_rt_match <- multiplot(p1,p2,cols = 2)
dev.off()
setwd(curDir)
multiplot(p1,p2,cols = 2)
```

```{r plot_categ_RT_val_data,echo=FALSE,warning=FALSE, message=FALSE}
RTdata1 <- df.C1.V.sum_rt_acc_l %>%
  select(Subject,Task,Morality,Identity,RT) %>% 
  filter(Task == "Val"& Morality == 'Good')
RTdata1$Identity <- factor(RTdata1$Identity,levels = c("Self","Other"))
p1 <- ggplot(data = RTdata1, aes(y = RT, x = Identity,fill = Identity)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0)) +
  geom_point(aes(y = RT,color = Identity), position = position_jitter(width = .15), size = .5) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  guides(fill = FALSE) +guides(color = FALSE)+
  theme_bw() +
  raincloud_theme+scale_y_continuous(limits = c(300,900))+labs(x = "Good",y = "Reaction times(ms)")
RTdata2 <- df.C1.V.sum_rt_acc_l %>%
  select(Subject,Task,Morality,Identity,RT) %>% 
  filter(Task == "Val"& Morality == 'Bad')
RTdata2$Identity <- factor(RTdata2$Identity,levels = c("Self","Other"))
p2 <- ggplot(data = RTdata2, aes(y = RT, x = Identity,fill = Identity)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0)) +
  geom_point(aes(y = RT,color = Identity), position = position_jitter(width = .15), size = .5) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  guides(fill = FALSE) +guides(color = FALSE)+
  theme_bw() +
  raincloud_theme+scale_y_continuous(limits = c(300,900),breaks = NULL)+labs(x = "Bad",y = "")

setwd(traDir)
tiff("p_rt_match.tiff", width = 9, height = 6, units = 'in', res = 300)
p_rt_match <- multiplot(p1,p2,cols = 2)
dev.off()
setwd(curDir)
multiplot(p1,p2,cols = 2)
```

```{r plot_categ_RT_Id_data,echo=FALSE,warning=FALSE, message=FALSE}
RTdata1 <- df.C1.V.sum_rt_acc_l %>%
  select(Subject,Task,Morality,Identity,RT) %>% 
  filter(Task == "Id"& Morality == 'Good')
RTdata1$Identity <- factor(RTdata1$Identity,levels = c("Self","Other"))
p1 <- ggplot(data = RTdata1, aes(y = RT, x = Identity,fill = Identity)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0)) +
  geom_point(aes(y = RT,color = Identity), position = position_jitter(width = .15), size = .5) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  guides(fill = FALSE) +guides(color = FALSE)+
  theme_bw() +
  raincloud_theme+scale_y_continuous(limits = c(300,900))+labs(x = "Good",y = "Reaction times(ms)")
                    #+ coord_cartesian(ylim=c(300,900))

#y_scale <- scale_y_continuous(limits = range(RTdata1$RT))
RTdata2 <- df.C1.V.sum_rt_acc_l %>%
  select(Subject,Task,Morality,Identity,RT) %>% 
  filter(Task == "Id"& Morality == 'Bad')
RTdata2$Identity <- factor(RTdata2$Identity,levels = c("Self","Other"))
p2 <- ggplot(data = RTdata2, aes(y = RT, x = Identity,fill = Identity)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0)) +
  geom_point(aes(y = RT,color = Identity), position = position_jitter(width = .15), size = .5) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  guides(fill = FALSE) +guides(color = FALSE)+
  theme_bw() +
  raincloud_theme+scale_y_continuous(limits = c(300,900),breaks = NULL)+labs(x = "Bad",y = "")

setwd(traDir)
tiff("p_rt_match.tiff", width = 9, height = 6, units = 'in', res = 300)
p_rt_match <- multiplot(p1,p2,cols = 2)
dev.off()
setwd(curDir)
multiplot(p1,p2,cols = 2)
```